### 2024-04-19 other/datasheet/Datasheet.qmd
Added datasheet.

---
title: "Datasheet"
author: 
  - Alaina Hu
date: today
date-format: long
format: pdf
number-sections: true
---

# Motivation

1. For what purpose was the dataset created? Was there a specific task in mind? Was there a specific gap that needed to be filled? Please provide a description.

The dataset was created to enable analysis of attitudes towards women and economic development on gender inequality. We were able to merge and create a dataset with the country, attitudes on women's income, attitudes on women's ability to be a political leader, attitudes justifying intimate-partner violence, and GDP along with the gender gap index score.

2. Who created the dataset (for example, which team, research group) and on behalf of which entity (for example, company, institution, organization)?

Alaina Hu, while learning at the University of Toronto

3. Who funded the creation of the dataset? If there is an associated grant, please provide the name of the grantor and the grant name and number.

No direct funding was received for this project.

4. Any other comments?

No.

# Composition

1. What do the instances that comprise the dataset represent (for example, documents, photos, people, countries)? Are there multiple types of instances (for example, movies, users, and ratings; people and interactions between them; nodes and edges)? Please provide a description.

Each row of the main dataset is a country of the world, each column in the row hold values for the country name, country code, attitude on women's income, attitude on women's ability to be a political leader, attitude justifying intimate partner violence, GDP, and gender gap index in that country

2. How many instances are there in total (of each type, if appropriate)?

68 rows with 7 variables

3. Does the dataset contain all possible instances or is it a sample (not necessarily random) of instances from a larger set? If the dataset is a sample, then what is the larger set? Is the sample representative of the larger set (for example, geographic coverage)? If so, please describe how this representativeness was validated/verified. If it is not representative of the larger set, please describe why not (for example, to cover a more diverse range of instances, because instances were withheld or unavailable).

No, there are only 68 countries. Data for many countries was not available in the dataset containing information on attitudes toward women. Attitude came from the OECD database, and the dataset only contains data for OECD countries and select non-OECD countries.

4. What data does each instance consist of? “Raw” data (for example, unprocessed text or images) or features? In either case, please provide a description.

Each instance consists of information such as percentages of population, monetary amounts so these variables are all positive integers or doubles.

5. Is there a label or target associated with each instance? If so, please provide a description.

Yes there is a unique key. The country name and country code differentiates the countries.

6. Is any information missing from individual instances? If so, please provide a description, explaining why this information is missing (for example, because it was unavailable). This does not include intentionally removed information, but might include, for example, redacted text.

All relevant information for this study is included in the data. No missing information for individual instances.

7. Are relationships between individual instances made explicit (for example, users’ movie ratings, social network links)? If so, please describe how these relationships are made explicit.

Yes, through the country name.

8. Are there recommended data splits (for example, training, development/validation, testing)? If so, please provide a description of these splits, explaining the rationale behind them.

No.

9. Are there any errors, sources of noise, or redundancies in the dataset? If so, please provide a description.

No.

10. Is the dataset self-contained, or does it link to or otherwise rely on external resources (for example, websites, tweets, other datasets)? If it links to or relies on external resources, a) are there guarantees that they will exist, and remain constant, over time; b) are there official archival versions of the complete dataset (that is, including the external resources as they existed at the time the dataset was created); c) are there any restrictions (for example, licenses, fees) associated with any of the external resources that might apply to a dataset consumer? Please provide descriptions of all external resources and any restrictions associated with them, as well as links or other access points, as appropriate.

Self-contained.

11. Does the dataset contain data that might be considered confidential (for example, data that is protected by legal privilege or by doctor-patient confidentiality, data that includes the content of individuals’ non-public communications)? If so, please provide a description.

No, all data were gathered from public sources.

12. Does the dataset contain data that, if viewed directly, might be offensive, insulting, threatening, or might otherwise cause anxiety? If so, please describe why.

No.

13. Does the dataset identify any sub-populations (for example, by age, gender)? If so, please describe how these subpopulations are identified and provide a description of their respective distributions within the dataset.

No.

14. Is it possible to identify individuals (that is, one or more natural persons), either directly or indirectly (that is, in combination with other data) from the dataset? If so, please describe how.

No.

15. Does the dataset contain data that might be considered sensitive in any way (for example, data that reveals race or ethnic origins, sexual orientations, religious beliefs, political opinions or union memberships, or locations; financial or health data; biometric or genetic data; forms of government identification, such as social security numbers; criminal history)? If so, please provide a description.

The dataset does not contains sensitive information.

16. Any other comments?

No.

# Collection process

1. How was the data associated with each instance acquired? Was the data directly observable (for example, raw text, movie ratings), reported by subjects (for example, survey responses), or indirectly inferred/derived from other data (for example, part-of-speech tags, model-based guesses for age or language)? If the data was reported by subjects or indirectly inferred/derived from other data, was the data validated/verified? If so, please describe how.

The data was gathered from the 2023 Global Gender Gap Report, OECD Gender, Institutions, and Development 2023 Database, and World Bank DataBank.

2. What mechanisms or procedures were used to collect the data (for example, hardware apparatuses or sensors, manual human curation, software programs, software APIs)? How were these mechanisms or procedures validated?

Loading into R. All data was provided in the form of a csv file or an excel file.

3. If the dataset is a sample from a larger set, what was the sampling strategy (for example, deterministic, probabilistic with specific sampling probabilities)?

The dataset is a sample of 68 countries of the world. The sample is selectively biased since not all countries had information reported in certain datasets, so analysis was not possible for those countries in this research.

4. Who was involved in the data collection process (for example, students, crowdworkers, contractors) and how were they compensated (for example, how much were crowdworkers paid)?

OECD, World Economic Forum, and World Bank

5. Over what timeframe was the data collected? Does this timeframe match the creation timeframe of the data associated with the instances (for example, recent crawl of old news articles)? If not, please describe the timeframe in which the data associated with the instances was created.

Gender gap index scores come from the 2023 Global Gender Gap Report. Attitude data comes from the 2023 OECD Gender, Institutions, and Development 2023 Database. GDP data comes from 2022 GDP per capita data from the World Bank.

6. Were any ethical review processes conducted (for example, by an institutional review board)? If so, please provide a description of these review processes, including the outcomes, as well as a link or other access point to any supporting documentation.

No.

7. Did you collect the data from the individuals in question directly, or obtain it via third parties or other sources (for example, websites)?

Third parties.

8. Were the individuals in question notified about the data collection? If so, please describe (or show with screenshots or other information) how notice was provided, and provide a link or other access point to, or otherwise reproduce, the exact language of the notification itself.

No.

9. Did the individuals in question consent to the collection and use of their data? If so, please describe (or show with screenshots or other information) how consent was requested and provided, and provide a link or other access point to, or otherwise reproduce, the exact language to which the individuals consented.

No, but individuals consented to the organizations using their data.

10. If consent was obtained, were the consenting individuals provided with a mechanism to revoke their consent in the future or for certain uses? If so, please provide a description, as well as a link or other access point to the mechanism (if appropriate).

Consent was not obtained.

11. Has an analysis of the potential impact of the dataset and its use on data subjects (for example, a data protection impact analysis) been conducted? If so, please provide a description of this analysis, including the outcomes, as well as a link or other access point to any supporting documentation.

No.

12. Any other comments?

No.

# Preprocessing/cleaning/labeling

1. Was any preprocessing/cleaning/labeling of the data done (for example, discretization or bucketing, tokenization, part-of-speech tagging, SIFT feature extraction, removal of instances, processing of missing values)? If so, please provide a description. If not, you may skip the remaining questions in this section.

Yes cleaning of the data was done.

2. Was the “raw” data saved in addition to the preprocessed/cleaned/labeled data (for example, to support unanticipated future uses)? If so, please provide a link or other access point to the “raw” data.

Yes, the raw data sets are in the data/raw_data folder, and scripts for the data cleaning are available in the scripts folder.

3. Is the software that was used to preprocess/clean/label the data available? If so, please provide a link or other access point.

R was used.

4. Any other comments?

No

# Uses

1. Has the dataset been used for any tasks already? If so, please provide a description.

The exact analysis dataset has not been used for any tasks because it is a unique combination of multiple data sets.

2. Is there a repository that links to any or all papers or systems that use the dataset? If so, please provide a link or other access point.
No

3. What (other) tasks could the dataset be used for?

Other variables would need to be added to the dataset.

4. Is there anything about the composition of the dataset or the way it was collected and preprocessed/cleaned/labeled that might impact future uses? For example, is there anything that a dataset consumer might need to know to avoid uses that could result in unfair treatment of individuals or groups (for example, stereotyping, quality of service issues) or other risks or harms (for example, legal risks, financial harms)? If so, please provide a description. Is there anything a dataset consumer could do to mitigate these risks or harms?

No.

5. Are there tasks for which the dataset should not be used? If so, please provide a description.

No.

6. Any other comments?

No.

# Distribution

1. Will the dataset be distributed to third parties outside of the entity (for example, company, institution, organization) on behalf of which the dataset was created? If so, please provide a description.

The dataset is available through GitHub. It is in the data/analysis_data folder of the repo.

2. How will the dataset be distributed (for example, tarball on website, API, GitHub)? Does the dataset have a digital object identifier (DOI)?

GitHub.

3. When will the dataset be distributed?

The dataset is available now.

4. Will the dataset be distributed under a copyright or other intellectual property (IP) license, and/or under applicable terms of use (ToU)? If so, please describe this license and/ or ToU, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms or ToU, as well as any fees associated with these restrictions.

No. MIT license.

5. Have any third parties imposed IP-based or other restrictions on the data associated with the instances? If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms, as well as any fees associated with these restrictions.

None that are known.

6. Do any export controls or other regulatory restrictions apply to the dataset or to individual instances? If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any supporting documentation.

None that are known.

7. Any other comments?

No.

# Maintenance

1. Who will be supporting/hosting/maintaining the dataset?

Alaina Hu

2. How can the owner/curator/manager of the dataset be contacted (for example, email address)?

alaina.hu@mail.utoronto.ca

3. Is there an erratum? If so, please provide a link or other access point.

No.

4. Will the dataset be updated (for example, to correct labeling errors, add new instances, delete instances)? If so, please describe how often, by whom, and how updates will be communicated to dataset consumers (for example, mailing list, GitHub)?

No, the dataset will not be updated.

5. If the dataset relates to people, are there applicable limits on the retention of the data associated with the instances (for example, were the individuals in question told that their data would be retained for a fixed period of time and then deleted)? If so, please describe these limits and explain how they will be enforced.

No.

6. Will older versions of the dataset continue to be supported/hosted/maintained? If so, please describe how. If not, please describe how its obsolescence will be communicated to dataset consumers.

The dataset has not been updated and will not need to be updated. It is available through GitHub.

7. If others want to extend/augment/build on/contribute to the dataset, is there a mechanism for them to do so? If so, please provide a description. Will these contributions be validated/verified? If so, please describe how. If not, why not? Is there a process for communicating/distributing these contributions to dataset consumers? If so, please provide a description.

Pull request on GitHub.

8. Any other comments?

No